---
title: "Fredericton Open Data Prep"
author: "Ryan Porter"
date: "September 27, 2017"
output: html_document
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), 'index.html')) })
---

First get the data fresh from the site

```{r}
if(!file.exists('bylaw.csv')){
  download.file('http://data.fredericton.ca/sites/default/files/by-law.csv', 'bylaw.csv')
  
}
curr_date <- date() 
bylaw <- read.csv('bylaw.csv', na.strings = c("NA", "NULL"))

```
This data was collected at `r curr_date`.

I think what I want to know is what kind of infraction is most common in which ward. A map of the wards is available [here](http://www.fredericton.ca/sites/default/files/pdf/2016may11wardmap.pdf). The first step will be cleaning the data a little bit. I will likely ask a couple other questions of the data later on.

```{r, fig.align = "left", fig.height = 8, fig.width = 15}
has_ward <- !is.na(bylaw$Ward)
bylaw.ward <- bylaw[has_ward,]

options(warn = -1)
suppressMessages(library(plotly))
p = plot_ly(x=bylaw.ward$Ward, type = "histogram", color=bylaw.ward$ComplaintCategory)
plotlink = layout(p, margin = 0)
plotlink

```

After attending the Open Data Atlantic fall session I've had some time to think about the data set and I have became more curious about the multi-entry reports (ie reports that share the same id). I am interested to find out if there is any additional information to find out from these entries. This will require data about each CaseID as there are a fair ammount of duplicates within the set (which I think I failed to take into account earlier).

```{r}
library(foreach)

get_CaseID_Data <- function(data, ID){
  tmp <- subset(data, CaseID == as.character(ID))
  num_entries <- nrow(tmp)
  dates <-tmp$SubDate
  ward <- tmp$Ward
  catagory <- tmp$ComplaintCategory
  return(list(ID, num_entries, dates, ward, catagory))
}


CaseID.data <- foreach(id=levels(bylaw$CaseID)) %do% get_CaseID_Data(bylaw, id)

```