---
title: "Fredericton Open Data Prep"
author: "Ryan Porter"
date: "September 27, 2017"
output: html_document
knit: (function(inputFile, encoding) { 
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(dirname(inputFile), 'index.html')) })
---

First get the data fresh from the site

```{r}
library(dplyr)
if(!file.exists('bylaw.csv')){
  download.file('http://data.fredericton.ca/sites/default/files/by-law.csv', 'bylaw.csv')
  
}
curr_date <- date() 
bylaw <- read.csv('bylaw.csv', na.strings = c("NA", "NULL"))
bylaw <- mutate(bylaw, SubDate = as.Date(SubDate))
```
This analysis was last knit on `r curr_date`.

I think what I want to know is what kind of infraction is most common in which ward. A map of the wards is available [here](http://www.fredericton.ca/sites/default/files/pdf/2016may11wardmap.pdf). The first step will be cleaning the data a little bit. I will likely ask a couple other questions of the data later on.

```{r, fig.align = "left", fig.height = 8, fig.width = 15}
has_ward <- !is.na(bylaw$Ward)
bylaw.ward <- bylaw[has_ward,]

options(warn = -1)
suppressMessages(library(plotly))
p = plot_ly(x=bylaw.ward$Ward, type = "histogram", color=bylaw.ward$ComplaintCategory)
plotlink = layout(p, margin = 0)
plotlink

```

After attending the Open Data Atlantic fall session I've had some time to think about the data set and I have became more curious about the multi-entry reports (ie reports that share the same id). I am interested to find out if there is any additional information to find out from these entries. This will require data about each CaseID as there are a fair ammount of duplicates within the set (which I think I failed to take into account earlier).

```{r, fig.width = 15}
library(foreach)
library(jsonlite)
library(dplyr)

get_CaseID_Data <- function(data, ID){
  # for use in a foreach loop.
  # gets data for rows matching ID in in the data (tailored for fred bylaw data)
  # ward and catagory only grab the value of the first entry though this is based on an asumtion on my part.
  
  tmp <- subset(data, CaseID == as.character(ID))
  num_entries <- nrow(tmp)
  dates <-tmp$SubDate
  ward <- tmp$Ward[1]
  catagory <- tmp$ComplaintCategory[1]
  
  return(list(ID=ID,
              num_entries=num_entries,
              date=dates,
              ward=as.character(ward), 
              type=as.character(catagory)
              ))
}

# collect data using above function and combine into a data.frame object
CaseID.data <- data.frame(foreach(id=levels(bylaw$CaseID), .combine = rbind ) %do% get_CaseID_Data(bylaw, id)) %>%
  mutate(ID = unlist(ID), num_entries=unlist(num_entries), ward=unlist(ward), type=unlist(type)) 

CaseID.data.json <- toJSON(CaseID.data, pretty = TRUE)
write(CaseID.data.json, file='byCaseID.json')
```
With that data gathered I can now look at posts by CaseID in the new dataframe. I have stored it as a JSON file because of the rows with multiple date values would not be compatible with csv. Next I want to see which types of reports have multiple values and which wards are inflated on the previous graph because of multiple reports.

```{r, fig.width = 14}

options(warn = -1)
suppressMessages(library(plotly))

font = list(
  family = "Courier New, monospace",
  size = 18,
  color = "#7f7f7f"
)
ylab = list(
  title = "number of unique IDs",
  titlefont = font
)
xlab = list(
  title = "number of posts per ID"
)

with(CaseID.data, plot_ly(x=num_entries, type = "histogram", color=type)) %>% 
  layout(xaxis = xlab, yaxis=ylab)

with(CaseID.data, plot_ly(x=num_entries, type = "histogram", color=ward)) %>%
  layout(xaxis = xlab, yaxis=ylab)

```

Looking at these charts it doesnt look like the number of posts in a ward should really throw the proportions off by all that much (each ward seems to have a similar proportion of multi entry IDs aside from ward 1). I may do further analysis on this later. 
